Scaling rates of true polar wander in convecting planets and moons

### 2017-05-01 tpw_rate.tex
Formatting.

The preceding results clarify the complex relationships that determine the rates of TPW for a convecting planet.
As mantle convection redistributes mass in the planet's interior, the spin axis moves around to stay aligned with the principal axes of the convective moment of inertia. 
There is a constant competition between growth of the mismatch angle $\theta$ through convective perturbations
 and its relaxation through TPW.
\citet{goldreich1969some} envisioned an analogy of beetles crawling on the surface of the globe, with
the spin axis trying to keep up with the instantaneous figure axis set by the beetles.
Our analysis begins to answer the questions ``how big are the beetles?'' and ``how fast are they crawling?’’. We find that the most important parameters are $m$, which acts as the brakes on the system, and the mismatch angle $\theta$. Surprisingly, the direct dependence of $\dot{W}$ on $Ra$ is weak (see Equation (44)), but the indirect influence of $Ra$ on the misalignment angle $\theta$ is quite strong.

Indeed, we can identify two endmember behaviors of Equation~\eqref{eq:simplest_milankovitch_defer_scaling}.
When convection is not sufficiently chaotic to create a large $\theta$ we are in the regime where the planet's rotation axis closely tracks that of the convective moment.
This is the regime considered in \citet{steinberger1997changes}, \citet{roberts2007cause}, and \citet{zhong2007supercontinent}, and can be considered the ``slow TPW'' regime.
When convection is more chaotic, however, there may be large excursions in $\theta$, 
which are driven by large changes in the orientation of the principal axes when $(\lambda_i - \lambda_j)$ is small.  
In this case a dramatic increase in the rate of TPW is possible. We refer to this as the `` fast TPW’’. Inertial interchanges occur when the maximum and intermediate axis are swapped and this can lead to a large change in 
 $\theta$ \citep{kirschvink1997evidence}.  Increasing $\theta$ to $\pi/2$ does not necessarily require fast TPW, but our numerical simulations show that rapid motion often occurs because $\theta$ does not remain close to $\pi/2$.  Moreover, the inertial interchange is only  a special case of the large $\theta$, ``fast TPW'' regime. We expect large $\theta$ and ``fast TPW’’ to be more prevalent in the early Earth, when the mantle was presumably hotter and less viscous (i.e. higher $Ra$). 
More vigorous convection increases the rate of change of density anomalies  and  makes
$\theta(t)$ less stable. Large (and potentially fast) TPW events should have been more frequent.

>>>>

The preceding results clarify the complex relationships that determine the rates of TPW for a convecting planet.
As mantle convection redistributes mass in the planet's interior, the spin axis moves around to stay aligned with the principal axes of the convective moment of inertia. 
There is a constant competition between growth of the mismatch angle $\theta$ through convective perturbations
 and its relaxation through TPW.
\citet{goldreich1969some} envisioned an analogy of beetles crawling on the surface of the globe, with
the spin axis trying to keep up with the instantaneous figure axis set by the beetles.
Our analysis begins to answer the questions ``how big are the beetles?'' and ``how fast are they crawling?''. We find that the most important parameters are $m$, which acts as the brakes on the system, and the mismatch angle $\theta$. Surprisingly, the direct dependence of $\dot{W}$ on $Ra$ is weak (see Equation (44)), but the indirect influence of $Ra$ on the misalignment angle $\theta$ is quite strong.

Indeed, we can identify two endmember behaviors of Equation~\eqref{eq:simplest_milankovitch_defer_scaling}.
When convection is not sufficiently chaotic to create a large $\theta$ we are in the regime where the planet's rotation axis closely tracks that of the convective moment.
This is the regime considered in \citet{steinberger1997changes}, \citet{roberts2007cause}, and \citet{zhong2007supercontinent}, and can be considered the ``slow TPW'' regime.
When convection is more chaotic, however, there may be large excursions in $\theta$, 
which are driven by large changes in the orientation of the principal axes when $(\lambda_i - \lambda_j)$ is small.  
In this case a dramatic increase in the rate of TPW is possible. We refer to this as the ``fast TPW''. Inertial interchanges occur when the maximum and intermediate axis are swapped and this can lead to a large change in 
 $\theta$ \citep{kirschvink1997evidence}.  Increasing $\theta$ to $\pi/2$ does not necessarily require fast TPW, but our numerical simulations show that rapid motion often occurs because $\theta$ does not remain close to $\pi/2$.  Moreover, the inertial interchange is only  a special case of the large $\theta$, ``fast TPW'' regime. We expect large $\theta$ and ``fast TPW'' to be more prevalent in the early Earth, when the mantle was presumably hotter and less viscous (i.e. higher $Ra$). 
More vigorous convection increases the rate of change of density anomalies  and  makes
$\theta(t)$ less stable. Large (and potentially fast) TPW events should have been more frequent.
